---
editor: 
  markdown: 
    wrap: 72
---

# Introduction

In this project I want to find out which variable influences the quality
of wine the most and if there are any correlations between the
variables. Another thing that I would like find is which wine in this
data set it the best wine because there are multiple wines that have an 8
but which one would be the best and why?

#### Meaning of the variables

variables:

1 - fixed acidity (tartaric acid - g / dm\^3) Most acids involved with
wine or fixed or nonvolatile

2 - volatileacidity (acetic acid - g /dm\^3) The amount of acetic acid
in wine, which at too high of levels can lead to an unpleasant, vinegar
taste 3 -citric acid (g / dm\^3)Found in small quantities, citric acid
can add 'freshness' and flavor to wines

4 - residual sugar (g / dm\^3) The amount of sugar remaining after
fermentation stops, it's rare to find wines with less than 1 gram/liter
and wines with greater than 45 grams/liter are considered sweet

5 -chlorides (sodium chloride - g / dm\^3 The amount of salt in the wine

6 - free sulfur dioxide (mg /dm\^3) The free form of SO2 exists in
equilibrium between molecular SO2 (as a dissolved gas) and bisulfite
ion; it prevents microbial growth and the oxidation of wine

7 - totalsulfur dioxide (mg / dm\^3) Amount of free and bound forms of
S02; in low concentrations, SO2 is mostly undetectable in wine, but at
free SO2 concentrations over 50 ppm, SO2 becomes evident in the nose and
taste of wine

8 - density (g / cm\^3)The density of water is close to that of water
depending on the percent alcohol and sugar content

9 - pH Describes how acidic or basic a wine is on a scale from 0 (very
acidic) to 14 (very basic); most wines are between 3-4 on the pH scale

10 -sulphates (potassium sulphate - g /dm3) A wine additive which can
contribute to sulfur dioxide gas (S02) levels, wich acts as an
antimicrobial and antioxidant

11 - alcohol (% by volume) The percent alcohol content of the wine
Output variable (based on sensory data):

12 - quality (score between 0 and 10)

### Data understanding and cleaning

The first step is the libraries. At the beginning of this project I
added the ones that I normally add and during my project I added some
more libraries that I needed for some of my code.

```{r message=FALSE, warning=FALSE}
library("ggplot2")
library("dplyr")
library("gridExtra")
library(GGally)
library(memisc)
library(pander)
library(corrplot)
library(fastDummies)
```

The next step is reading the CSV file.

```{r message=FALSE, warning=FALSE}
Winee <- read.csv('wineQualityReds.csv')

```

In this step I added a rating to the data to get a quick view on the
rating.

```{r}
Winee$rating <- ifelse(Winee$quality < 6, 'bad', ifelse(
  Winee$quality < 7, 'average', 'good'))

Winee$rating <- ordered(Winee$rating,
                       levels = c('bad', 'average', 'good'))
```

I used the str function to see the types of data and to see of I need to
change anything.

```{r message=FALSE, warning=FALSE, results='hide'}
str(Winee)
```

My next step is looking at the summary to see if there would be a lot of
outliers or not.

```{r message=FALSE, warning=FALSE, packages, results='hide'}
summary(Winee)
```

### Analysis

The summary was not very clear for me so I decided to make some
histograms to get an even better look at the data.

```{r}
ggplot(Winee) +
  aes(x = fixed.acidity) +
  geom_histogram(bins = 30L, fill = "#722F37") +
  theme_minimal()
```

```{r}
ggplot(Winee) +
  aes(x = volatile.acidity) +
  geom_histogram(bins = 30L, fill = "#722F37") +
  theme_minimal()
```

```{r}
ggplot(Winee) +
  aes(x = citric.acid) +
  geom_histogram(bins = 30L, fill = "#722F37") +
  theme_minimal()
```

```{r}
ggplot(Winee) +
  aes(x = residual.sugar) +
  geom_histogram(bins = 30L, fill = "#722F37") +
  theme_minimal()
```

```{r}
ggplot(Winee) +
  aes(x = chlorides) +
  geom_histogram(bins = 30L, fill = "#722F37") +
  theme_minimal()
```

```{r}
ggplot(Winee) +
  aes(x = free.sulfur.dioxide) +
  geom_histogram(bins = 30L, fill = "#722F37") +
  theme_minimal()
```

```{r}
ggplot(Winee) +
  aes(x = total.sulfur.dioxide) +
  geom_histogram(bins = 30L, fill = "#722F37") +
  theme_minimal()
```

```{r}
ggplot(Winee) +
  aes(x = density) +
  geom_histogram(bins = 30L, fill = "#722F37") +
  theme_minimal()
```

```{r}
ggplot(Winee) +
  aes(x = pH) +
  geom_histogram(bins = 30L, fill = "#722F37") +
  theme_minimal()
```

```{r}
ggplot(Winee) +
  aes(x = sulphates) +
  geom_histogram(bins = 30L, fill = "#722F37") +
  theme_minimal()
```

```{r}
ggplot(Winee) +
  aes(x = alcohol) +
  geom_histogram(bins = 30L, fill = "#722F37") +
  theme_minimal()
```

It is important to find correlations so I created a data frame with all
the correlations.

```{r message=FALSE, warning=FALSE, results='hide'}
mcor1<-(cor(Winee$fixed.acidity, Winee[c( "fixed.acidity", "volatile.acidity", "citric.acid", "residual.sugar", "chlorides", "free.sulfur.dioxide", "total.sulfur.dioxide", "density", "pH", "sulphates", "alcohol")]))
mcor2<-(cor(Winee$volatile.acidity, Winee[c( "fixed.acidity", "volatile.acidity", "citric.acid", "residual.sugar", "chlorides", "free.sulfur.dioxide", "total.sulfur.dioxide", "density", "pH", "sulphates", "alcohol")]))
mcor3<-(cor(Winee$citric.acid, Winee[c( "fixed.acidity", "volatile.acidity", "citric.acid", "residual.sugar", "chlorides", "free.sulfur.dioxide", "total.sulfur.dioxide", "density", "pH", "sulphates", "alcohol")]))
mcor4<-(cor(Winee$residual.sugar, Winee[c( "fixed.acidity", "volatile.acidity", "citric.acid", "residual.sugar", "chlorides", "free.sulfur.dioxide", "total.sulfur.dioxide", "density", "pH", "sulphates", "alcohol")]))
mcor5<-(cor(Winee$chlorides, Winee[c( "fixed.acidity", "volatile.acidity", "citric.acid", "residual.sugar", "chlorides", "free.sulfur.dioxide", "total.sulfur.dioxide", "density", "pH", "sulphates", "alcohol")]))
mcor6<-(cor(Winee$free.sulfur.dioxide, Winee[c( "fixed.acidity", "volatile.acidity", "citric.acid", "residual.sugar", "chlorides", "free.sulfur.dioxide", "total.sulfur.dioxide", "density", "pH", "sulphates", "alcohol")]))
mcor7<-(cor(Winee$total.sulfur.dioxide, Winee[c( "fixed.acidity", "volatile.acidity", "citric.acid", "residual.sugar", "chlorides", "free.sulfur.dioxide", "total.sulfur.dioxide", "density", "pH", "sulphates", "alcohol")]))
mcor8<-(cor(Winee$density, Winee[c( "fixed.acidity", "volatile.acidity", "citric.acid", "residual.sugar", "chlorides", "free.sulfur.dioxide", "total.sulfur.dioxide", "density", "pH", "sulphates", "alcohol")]))
mcor9<-(cor(Winee$pH, Winee[c( "fixed.acidity", "volatile.acidity", "citric.acid", "residual.sugar", "chlorides", "free.sulfur.dioxide", "total.sulfur.dioxide", "density", "pH", "sulphates", "alcohol")]))
mcor10<-(cor(Winee$sulphates, Winee[c( "fixed.acidity", "volatile.acidity", "citric.acid", "residual.sugar", "chlorides", "free.sulfur.dioxide", "total.sulfur.dioxide", "density", "pH", "sulphates", "alcohol")]))
mcor11<-(cor(Winee$alcohol, Winee[c( "fixed.acidity", "volatile.acidity", "citric.acid", "residual.sugar", "chlorides", "free.sulfur.dioxide", "total.sulfur.dioxide", "density", "pH", "sulphates", "alcohol")]))
```

```{r message=FALSE, warning=FALSE, results='hide'}
correlation<-rbind.data.frame(mcor1,mcor2,mcor3,mcor4,mcor5,mcor6,mcor7,mcor8,mcor9,mcor10,mcor11, deparse.level = 1, make.row.names = TRUE, stringsAsFactors = FALSE, factor.exclude = TRUE )
```

\tiny

my first correlation data frame is too messy for me so I decided to
round the numbers and to add names on the left to create a data frame
that looks a bit more like a matrix.

```{r results='hide', message=FALSE, warning=FALSE}
correlations.rounded<- round(correlation, 2)
print(correlations.rounded)
```

```{r message=FALSE, warning=FALSE}
factor<- c("fixed.acidity", "volatile,acidity", "citric.acid", "residual.sugar", "chlorides", "free.sulfur.dioxide", "total.sulfur.dioxide", "density", "pH", "sulphates", "alcohol")

correlations.rounded$factor <- factor

correlations.rounded<- correlations.rounded %>% relocate(factor, .before = fixed.acidity)

print(correlations.rounded)

```

to really show the correlations I decided to create boxplots and see
where the correlations are and which variables have the most influence.

```{r message=FALSE, warning=FALSE}
ggplot(data = Winee, aes(x = quality, y = fixed.acidity)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = '#722F37') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)       
```

```{r message=FALSE, warning=FALSE}
ggplot(data = Winee, aes(x = quality, y = volatile.acidity)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = '#722F37') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)
```

```{r message=FALSE, warning=FALSE}
ggplot(data = Winee, aes(x = quality, y = residual.sugar)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = '#722F37') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)
```

```{r message=FALSE, warning=FALSE}
ggplot(data = Winee, aes(x = quality, y = chlorides)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = '#722F37') +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)
```

```{r message=FALSE, warning=FALSE}
ggplot(data=Winee, aes(x=quality, y=free.sulfur.dioxide)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = '#722F37') +
  scale_y_continuous(lim = c(0,40)) +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)
```

```{r message=FALSE, warning=FALSE}
ggplot(data=Winee, aes(x=quality, y=total.sulfur.dioxide)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = '#722F37') +
  scale_y_continuous(lim = c(0,40)) +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)
```

```{r message=FALSE, warning=FALSE}
ggplot(data=Winee, aes(x=quality, y=density)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = '#722F37') +
  scale_y_continuous(lim = c(0.95,1.05)) +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)
```

```{r message=FALSE, warning=FALSE}
ggplot(data=Winee, aes(x=quality, y=pH)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .5,color = '#722F37') +
  scale_y_continuous(lim = c(2,5)) +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)
```

```{r message=FALSE, warning=FALSE}
ggplot(data=Winee, aes(x=quality, y=alcohol)) +
  geom_jitter( alpha = .3) +
  geom_boxplot(alpha = .10,color = '#722F37') +
  scale_y_continuous(lim = c(8,15)) +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)
```

what I can see is that a higher alcohol percentage shows a higher
quality. but there are a lot of outliers so it might be that there are
also other things that contribute to this. So my next step is to create
a simple linear model.

```{r message=FALSE, warning=FALSE}
alcoholQualityLinearModel <- lm(as.numeric(quality) ~ alcohol,
                       data = Winee)
summary(alcoholQualityLinearModel)

```

```{r message=FALSE, warning=FALSE}
simple_cor_test <- function(x, y) {
  return(cor.test(x, as.numeric(y))$estimate)
}

correlations <- c(
  simple_cor_test(Winee$fixed.acidity, Winee$quality),
  simple_cor_test(Winee$volatile.acidity, Winee$quality),
  simple_cor_test(Winee$citric.acid, Winee$quality),
  simple_cor_test(log10(Winee$residual.sugar), Winee$quality),
  simple_cor_test(log10(Winee$chlorides), Winee$quality),
  simple_cor_test(Winee$free.sulfur.dioxide, Winee$quality),
  simple_cor_test(Winee$total.sulfur.dioxide, Winee$quality),
  simple_cor_test(Winee$density, Winee$quality),
  simple_cor_test(Winee$pH, Winee$quality),
  simple_cor_test(log10(Winee$sulphates), Winee$quality),
  simple_cor_test(Winee$alcohol, Winee$quality))
names(correlations) <- c('fixed.acidity', 'volatile.acidity', 'citric.acid',
                         'log10.residual.sugar',
                         'log10.chlordies', 'free.sulfur.dioxide',
                         'total.sulfur.dioxide', 'density', 'pH',
                         'log10.sulphates', 'alcohol')

correlations
```

after this I can now see that the following variables have a high
correlation with the quality of wine: Alcohol

Sulphates(log10)

Volatile Acidity

Citric Acid

### results

1.  Fixed Acidity seems to have almost no effect on quality because the
    correlation is very low.

2.  Volatile Acidity seems to have a negative correlation with the
    quality.

3.  Better wines seem to have higher concentration of Citric Acid.

4.  Better wines seem to have higher alcohol percentages. But when I
    created a linear model around it, I saw from the R squared value
    that alcohol by itself only contributes like 20% on the variance of
    the quality. So there may be some other factors at play here. Even
    though it's a weak correlation, but lower percent of Chloride seems
    to produce better quality wines as can be seen in the graph.

5.  Better wines seem to have lower densities. But then again, this may
    be due to the higher alcohol content in them.

6.  Better wines seem to be more acidic.

7.  Residual sugar almost has no effect on the wine quality.

```{r message=FALSE, warning=FALSE}
df2<- Winee[,c('X', 'fixed.acidity', 'volatile.acidity', 'citric.acid', 'residual.sugar', 'chlorides', 'free.sulfur.dioxide', 'total.sulfur.dioxide', 'density', 'pH', 'sulphates', 'alcohol', 'quality')]
```

```{r message=FALSE, warning=FALSE}
TopWine<- filter(df2, quality == '8')

```

I can now see that there are 18 wines in my data set with an 8 and thus
are the best but now I want to see if my variables that I found before
are also coming back in these 18 wines and I want to see which of these
wines is the number one

```{r message=FALSE, warning=FALSE}
TopWine
```

the things that I look at are: - high concentration of citric
acid(441) - high alcohol %(1270) - low % of chlorides(1270) - low
dinsity(1270) - low PH (meaning more acidic)(441) (the numbers behind
indiciate the wine that scores best in this topic)

according to this analysis wine number 441 and wine number 1270 are the
best wines in this data set.

```{r message=FALSE, warning=FALSE}
TopWine.2<- filter(TopWine, X == '1270' | X == '441')
TopWine.2
```

### Reflection
During this project I learned a lot more about writing code and creating visuals. I had a few moments where is was totally stuck and kept getting errors this was very frustrating but it also helped me to find creative solutions and to get better at formulating questions in Google. In the beginning of this project I wanted to find out the quality of my own favourite wine but I came to the conclusion that finding these kind of variables about my own favourite wine is very difficult so I decided to look at what are the most important variables and started looking for the best wine in the data that I had. 


